---
title: "Uncertainty Analysis Using PEcAn"
author: 
  - "Aritra Dey"
  - "David LeBauer"
format:
  pdf:
    toc: true
    number-sections: true
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
---

# Introduction {#introduction}

In Demo 2 we will be looking at how PEcAn can use information about parameter uncertainty to perform four related analyses:

- **Ensemble Analysis**: Repeats model runs while sampling from parameter uncertainty to propagate parameter uncertainty through to model output and estimate uncertainty in model projections (LeBauer et al. 2013).
- **Sensitivity Analysis**: Repeats model runs to assess how changes in model parameters affect model outputs. Allows us to identify which parameters the model is most sensitive to.
- **Variance Decomposition**: Combines the one-at-a-time sensitivity results with parameter uncertainty to estimate which parameters contribute most to model uncertainty and help prioritize subsequent data collection (LeBauer et al. 2013).
- **Global Sensitivity Analysis**: Uses a Sobol design to partition output variance across major sources of uncertainty such as parameters, meteorology, initial conditions, and events. In contrast to one-at-a-time analysis, this global design can account for interactions among factors (Saltelli et al. 2008).

This demo shows how to run an uncertainty analysis workflow in PEcAn using an R-based Quarto notebook. It covers loading settings, configuring models, running simulations, and performing ensemble analysis, one-at-a-time sensitivity analysis, variance decomposition, and Sobol global sensitivity analysis. This programmatic approach complements the web-based PEcAn interface.

Dietze (2017) provides a comprehensive review of sensitivity and uncertainty analysis in ecosystem modeling.

**Context & modeling scenario:**

We simulate plant and ecosystem carbon balance (Net Primary Productivity and Net Ecosystem Exchange) at the AmeriFlux Niwot Ridge Forest site ([US‑NR1](https://ameriflux.lbl.gov/sites/siteinfo/US-NR1)) during the year 2004. We use SIPNET parameterized as a temperate conifer PFT and driven by AmeriFlux meteorology following the analysis in [Moore et al. (2007)](https://doi.org/10.1016/j.agrformet.2008.04.013). This notebook also provides a compact template that can be extended to more years, locations, and PFTs.

**What this notebook does:**

1. Configure a PEcAn workflow by loading and validating a `pecan.xml` settings file.
2. Run a set of ecosystem model simulations by writing model configuration files and then running SIPNET.
3. Quantify uncertainty using ensemble analysis, one-at-a-time sensitivity analysis, variance decomposition, and Sobol global sensitivity analysis.
4. Visualize results to identify important parameters and how they influence model variance.
5. Change configuration settings and re-run the workflow.

## Prerequisites

To run this notebook, you will need to install PEcAn and its dependencies, as well as download the SIPNET model binary.

### PEcAn packages and dependencies.

```
# Enable repository from pecanproject
options(repos = c(
  pecanproject = 'https://pecanproject.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
install.packages('PEcAn.all')
```

A valid `pecan.xml` configuration file. Start with the example at `pecan/documentation/tutorials/Demo_02_Uncertainty_Analysis/pecan.xml`.

### SIPNET v1.3.0

If you haven't already installed the SIPNET model, you can do so by running the following code. This will download the SIPNET binary to `demo_outdir/sipnet` and make it executable.

> Note: The `demo_outdir` directory will be created inside of your PEcAn installation, at `documentation/tutorials/Demo_02_Uncertainty_Analysis/demo_outdir/`. This directory will contain the SIPNET binary as well as the output generated by PEcAn in this demo.

```{r download-sipnet}
# Download and install SIPNET v1.3.0
source(
  here::here(
    "documentation/tutorials/Demo_1_Basic_Run/download_sipnet.R"
  )
)
```

> Note: You can find the most recent version of the SIPNET binary at: [SIPNET GitHub Releases](https://github.com/PecanProject/sipnet/releases), but this notebook is designed to work with SIPNET v1.3.0.

# Load PEcAn Packages

First, we need to load the PEcAn R packages. These packages provide all the functions we'll use to run the workflow.

```{r libraries}
# Load the PEcAn.all package, which includes all necessary PEcAn functionality
library("PEcAn.all")
```

# Load PEcAn Settings File

Use the XML settings file (`pecan.xml`) exactly as in the Demo 1 Basic Run tutorial. See the "Load PEcAn Settings File" section of Demo 1 for a more detailed walkthrough of fields and schema. In this tutorial we focus on settings relevant to this notebook and explain the additional options associated with uncertainty analysis.

## Settings

Example settings for this demo live at `pecan/documentation/tutorials/Demo_02_Uncertainty_Analysis/pecan.xml` and you can read more about the settings in the [PEcAn Documentation](https://pecanproject.github.io/pecan-documentation/develop/pecanXML.html#pecanXML), and sections focused on [ensemble](https://pecanproject.github.io/pecan-documentation/develop/xml-advanced.html#xml-ensemble) and [sensitivity analysis](https://pecanproject.github.io/pecan-documentation/develop/xml-advanced.html#xml-sensitivity-analysis) settings in particular.

Open that settings file and look at the ensemble and sensitivity analysis sections. You can modify these settings to change the number of ensemble members, the variable to analyze, and the sampling method. Some of the key settings in this demo include:

**Ensemble size**

The number of runs in the ensemble is set to **50** in this demo, but a larger ensemble size (100-5000) is often used in practice to better estimate the posterior distribution of the output.

**Output variable**

The output variable for the ensemble analysis is set to **NPP** (Net Primary Productivity). You can change this to other variables like NEE (Net Ecosystem Exchange), LAI (Leaf Area Index), ET (Evapotranspiration), etc. You can also specify multiple variables by providing a vector of variable names (e.g., `c("NPP", "NEE", "LAI")`) to analyze uncertainty across several ecosystem processes simultaneously.

**Sampling method**

The sampling method for generating parameter sets is set to **halton** in this demo. This tells PEcAn to sample using the [Halton sequence](https://en.wikipedia.org/wiki/Halton_sequence), which is a quasi-random sampling method that more efficiently explores parameter space than random sampling. Other options include Latin Hypercube (lhc), which is another quasi-random sequence, as well as uniform that draws random samples. A random sampler requires a larger ensemble size to adequately explore parameter space.

**Sensitivity analysis quantiles**

PEcAn's sensitivity analysis includes a handy shortcut that converts a specified standard deviation into its normal quantile equivalent. In the example pecan.xml, these are set to **-1, 1** (the median value, 0, occurs by default) which are converted internally to the 15.9th and 84.1th quantiles of the parameter distribution. You can add more quantiles to explore a wider range of parameter values: {-3, -2, -1, 1, 2, 3} is often used in practice. 

By working in quantiles relative to each parameter's distribution, the sensitivity and variance decomposition reflect sensitivity across the range of parameter values. This matches the PEcAn goal of propagating parameter uncertainty through to model output and then evaluating how much each parameter contributes to model uncertainty (LeBauer et al. 2013). Many one-at-a-time analyses use a fixed perturbation such as mean +/- 10%, but that does not capture uncertainty across the parameter distribution and therefore does not support the same variance decomposition workflow.

## Load the settings file

```{r load-settings}
settings_path <- here::here("documentation/tutorials/Demo_02_Uncertainty_Analysis/pecan.xml")

settings <- PEcAn.settings::read.settings(settings_path)

settings <- PEcAn.settings::prepare.settings(settings)
```

See Demo 1 Section 6 for details on what these functions do. Briefly, they read the XML file, convert it into an R list object that PEcAn can use, check that settings are valid, fill in defaults, and create the output directory.

# Write Model Configuration Files {#sec-write-configs}

This step generates the model-specific configuration files that will be used to run the ecosystem model. The process involves:

1. Disabling database write operations because we are not using a database
```{r disable-db-write}
settings$database <- NULL # Disable database operations for this demo
```
2. Generating SIPNET configuration files using the `runModule.run.write.configs()` function.
```{r write-configs}
settings <- PEcAn.workflow::runModule.run.write.configs(settings)
```

# Run Model Simulations

This section executes the SIPNET simulations and retrieves the results.

It uses the function `runModule_start_model_runs(settings)` to initiate the model runs using the configuration files generated in the previous step.

```{r run-model}
PEcAn.workflow::runModule_start_model_runs(settings)
```

The PEcAn workflow will take a longer time to complete than in Demo 1 because we have just asked for over a hundred model runs. Once the runs are complete we will continue.

# Fetch Model Outputs

Next we convert all of the model output from the previous run to a standard format that PEcAn can use for analysis. This is done using the `runModule.get.results()` function.

```{r get-model-results}
runModule.get.results(settings)
```

# Ensemble, Sensitivity, Variance Decomposition, and Global Sensitivity Analysis {#sec-run-uncertainty}

Next, we use the outputs from the previous step for two related workflows: an ensemble plus one-at-a-time sensitivity workflow that feeds the variance decomposition analysis, and a separate Sobol global sensitivity workflow.

**Ensemble Analysis**: Quantifies uncertainty in model predictions by running multiple simulations with parameters sampled from their uncertainty distributions. This generates probability distributions of model outputs and confidence intervals.

```{r run-ensemble-analysis}
runModule.run.ensemble.analysis(settings)
```

**Sensitivity Analysis**: Systematically varies individual parameters to assess their influence on model outputs. In this demo, these one-at-a-time runs are also the basis for the variance decomposition analysis that follows. This is useful for prioritizing parameters, but it does not estimate interactions among uncertain factors.

```{r run-sensitivity-analysis}
runModule.run.sensitivity.analysis(settings)
```

## Sobol Global Sensitivity Analysis

Sobol global sensitivity analysis is a separate variance-based workflow. Rather than using the one-at-a-time sensitivity results, it starts from a Sobol design and estimates first-order and total-order contributions directly from a global run ensemble. This is a variance-based global sensitivity analysis: it diagnoses how uncertainty in the inputs contributes to uncertainty in the model output, including interactions among factors (Dietze 2017; Saltelli et al. 2008). PEcAn computes these indices using `sensobol` (Puy et al. 2022). It is often helpful to run this analysis in a new output directory to clearly separate runs generated by the Sobol design.

> Note: In Sobol mode, `ensemble_size` means the base sample size `N`, not the expanded number of model runs.

```{r run-sobol-gsa, eval=FALSE}
gsa_settings <- settings
# Use a fresh output directory before running this block.

sobol_design <- PEcAn.uncertainty::generate_joint_ensemble_design(
  settings = gsa_settings,
  ensemble_size = 16,
  sobol = TRUE
)

gsa_settings <- PEcAn.workflow::runModule.run.write.configs(
  settings = gsa_settings,
  input_design = sobol_design$X
)
PEcAn.workflow::runModule_start_model_runs(gsa_settings)
runModule.get.results(gsa_settings)

sobol_indices <- PEcAn.uncertainty::compute_sobol_indices(
  outdir = gsa_settings$outdir,
  sobol_obj = sobol_design,
  var = "NPP"
)

sobol_indices
```

# PEcAn Outputs {#sec-outputs}

## Output Directory Structure

These are the key folders and files that will be created under the directory defined by `settings$outdir` (e.g., `demo_outdir` in the example). The file contents are described in the next section.

We discussed the output directory in Demo 1 (Basic Run), but now we have additional outputs from the ensemble, one-at-a-time sensitivity, variance decomposition, and Sobol global sensitivity analyses. Here we focus on those additional analysis products.

Note: The layout below is representative - the actual files in the out directory will depend
on workflow settings.

```
demo_outdir/
├── run/                                    # Configuration & execution metadata
│   ├── runs.txt                           # List of run IDs (SA and ensemble runs)
│   ├── ENS-*-*/                          # Ensemble run directories (e.g., ENS-00001--1/)
│   └── SA-*-*/                           # Sensitivity analysis run directories
├── out/                                   # Raw model outputs by run ID
│   └── <runid>/                          # E.g., daily or sub-daily SIPNET output files
├── ensemble.analysis.*.pdf                # Ensemble analysis plots
├── ensemble.output.*.Rdata               # Raw ensemble outputs
├── ensemble.samples.*.Rdata              # Parameter samples used for ensemble
├── ensemble.ts.*.Rdata                   # Time series data from ensemble
├── samples.Rdata                         # Parameter samples for both SA and ensemble
├── sobol.design.*.Rdata                  # Saved Sobol design metadata for post-processing
├── sobol.indices.*.Rdata                 # Saved Sobol first-order and total-order indices
├── sensitivity.output.*.Rdata            # SA model outputs
├── sensitivity.results.*.Rdata           # Processed SA results
├── sensitivity.samples.*.Rdata           # SA parameter samples
├── variance.decomposition.*.pdf          # Variance decomposition analysis
├── pft/                                  # Parameter (prior/posterior) files per PFT
│   └── temperate.coniferous/
└── sipnet                                # SIPNET binary (downloaded earlier)
```

### Model outputs and logs

- Standardized netCDF files (`[year].nc`) for analysis and visualization
- Raw model output (for SIPNET, e.g., `sipnet.out` per run)
- `logfile.txt` with model and workflow messages
- Note: `pft/` contains parameter files used in estimation; see the parameter-estimation tutorial (Demo 3) for details

# Understanding PEcAn Uncertainty Analysis Outputs

After running ensemble analysis, one-at-a-time sensitivity analysis, variance decomposition, and global sensitivity analysis, PEcAn generates several important outputs that help you understand model uncertainty and parameter sensitivity.

The `samples.Rdata` file contains the parameter values used in the sensitivity and ensemble runs. It stores two objects, `sa.samples` and `ensemble.samples`, which are the parameter values for the sensitivity analysis and ensemble runs, respectively.

## Ensemble Analysis Outputs

The ensemble analysis produces:

- **`ensemble.output.[RunID].[Variable].[StartYear].[EndYear].Rdata`**: Contains the `ensemble.output` object with model predictions for all ensemble members
- **`ensemble.analysis.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Histogram and boxplot of ensemble predictions
- **`ensemble.ts.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Time-series plot showing ensemble mean, median, and 95% confidence intervals

## Sensitivity Analysis Outputs

The sensitivity analysis generates:

- **`sensitivity.analysis.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Raw data points from univariate analyses with spline fits.
- **`sensitivity.output.[RunID].[Variable].[StartYear].[EndYear].Rdata`**: Model outputs corresponding to parameter variations.
- `sensitivity.analysis.[RunID].[Variable].[StartYear].[EndYear].pdf` shows the raw data points from univariate one-at-a-time analyses and spline fits through the points. _Open this file_ to determine which parameters are most and least sensitive.

## Variance Decomposition Outputs

The variance decomposition produces:

- **`variance.decomposition.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Three-column analysis showing:
  - Coefficient of variation (normalized posterior variance)
  - Elasticity (normalized sensitivity) 
  - Partial standard deviation of each parameter


## Sobol Global Sensitivity Outputs

The Sobol global sensitivity analysis generates:

- **`sobol.design.[RunID].[Variable].[StartYear].[EndYear].Rdata`**: The saved Sobol design metadata used to map PEcAn runs back to the Sobol estimator.
- **`sobol.indices.[RunID].[Variable].[StartYear].[EndYear].Rdata`**: The computed first-order and total-order Sobol indices for each uncertainty class in the design. First-order indices summarize the direct contribution of each factor, whereas total-order indices summarize the factor's overall contribution including interactions (Saltelli et al. 2008). PEcAn computes these indices using `sensobol` (Puy et al. 2022).

After running the Sobol workflow, you can inspect `sobol_indices` directly or load the saved `sobol.indices.*.Rdata` file to compare first-order and total-order contributions across uncertainty classes.

## Interpreting the Results

**Variance Decomposition Analysis:**

- Parameters are sorted by their contribution to model output uncertainty (the right column).
- Identify parameters that are:
  - Highly sensitive but low uncertainty.
  - Highly uncertain but low sensitivity.
  - Both sensitive and uncertain.
- Identify parameters that are both sensitive and uncertain for future constraint with data or expert knowledge.
- Potential gotchas:
  - Flat sensitivity curves: check that parameter values were correctly generated and read by the model.
  - Parameters with high uncertainty: consider revising priors.
  - Multi-modal or otherwise unexpected parameter distributions: check that parameter was specified correctly.

 **Choose the parameter that you think provides the most efficient means of reducing model uncertainty and propose how you might best reduce uncertainty in this process**. In making this choice remember that not all processes in models can be directly observed, and that the cost-per-sample for different measurements can vary tremendously (and thus the parameter you measure next is not always the one contributing the most to model variability). Also consider the role of parameter uncertainty versus model sensitivity in justifying your choice of what parameters to constrain.
  
# Visualize Uncertainty Analysis Results {#sec-visualize}

This section loads the results from the uncertainty analyses and generates plots directly in the notebook. This provides an immediate view of the ensemble time series, sensitivity plots, and variance decomposition.

## Ensemble Analysis Visualization

Here we visualize the results of the ensemble analysis. It shows the overall distribution of the model output and how the output and its uncertainty change over time.

This section reproduces the plots saved in the `ensemble.analysis/` folder in order to show the user how to access and visualize the results programmatically so that they can further investigate output and customize plots.

```{r visualize-ensemble-results}
# --- 1. Define Helper Variables ---
# Extract key variables from the settings object to simplify file path construction
# and plotting. This makes the code cleaner and easier to read.
variable <- settings$ensemble$variable
pft <- settings$pfts[[1]]
start.year <- lubridate::year(settings$run$start.date)
end.year <- lubridate::year(settings$run$end.date)
ensemble.id <- if (!is.null(settings$ensemble$ensemble.id)) settings$ensemble$ensemble.id else "NOENSEMBLEID"

# --- 2. Load and Plot Ensemble Output Distribution ---
# This section visualizes the distribution of the ensemble model runs.
# It generates a histogram and a boxplot to show the central tendency, spread,
# and shape of the output variable's distribution.

# Construct the path to the ensemble output file
ens_file <- PEcAn.uncertainty::ensemble.filename(
  settings,
  prefix = "ensemble.output",
  ensemble.id = ensemble.id,
  variable = variable,
  start.year = start.year,
  end.year = end.year
)

# Check if the file exists, then load and plot the data
if (file.exists(ens_file)) {
  ens_env <- new.env()
  load(ens_file, envir = ens_env)
  ens_data <- as.numeric(unlist(ens_env$ensemble.output))

  # Define units for plot labels
  units <- paste0(variable, " (", PEcAn.utils::mstmipvar(variable, silent = TRUE)$units, ")")

  # Create side-by-side histogram and boxplot
  par(mfrow = c(1, 2), mar = c(4, 4.8, 1, 2))
  hist(ens_data, xlab = units, main = "Ensemble Distribution", cex.lab = 1.4, col = "grey85")
  box(lwd = 2.2)
  boxplot(ens_data, ylab = units, main = "Ensemble Boxplot", col = "grey85", cex.lab = 1.5)
  box(lwd = 2.2)
  par(mfrow = c(1, 1))
} else {
  PEcAn.logger::logger.warn("Could not find ensemble output file:", ens_file)
}


# --- 3. Plot Ensemble Time Series ---
# This section visualizes the ensemble results over time, showing the mean,
# median, and 95% confidence interval of the model output.
ens_ts_data <- PEcAn.uncertainty::read.ensemble.ts(settings, variable = variable)
if (!is.null(ens_ts_data)) {
  PEcAn.uncertainty::ensemble.ts(ens_ts_data)
}
```

## Sensitivity and Variance Decomposition Visualization

This block visualizes the results of the sensitivity analysis. The plots show how sensitive the model output is to changes in each parameter and which parameters contribute most to the overall uncertainty.

```{r visualize-sensitivity-results}
# --- 1. Load Sensitivity Analysis Results ---
# This section loads the saved sensitivity analysis data, which contains the
# outputs needed to generate the sensitivity and variance decomposition plots.

# Construct the path to the sensitivity analysis results file
sensitivity.id <- if (!is.null(settings$sensitivity.analysis$ensemble.id)) {
  settings$sensitivity.analysis$ensemble.id
} else {
  "NOENSEMBLEID"
}
sens_file <- file.path(
  settings$outdir,
  paste0("sensitivity.results", ".", sensitivity.id, ".", variable, ".", start.year, ".", end.year, ".Rdata")
)

# Check if the file exists, then load the data and generate plots
if (file.exists(sens_file)) {
  sens_env <- new.env()
  load(sens_file, envir = sens_env)
  sensitivity.results <- sens_env$sensitivity.results

  # --- 2. Generate Sensitivity Plots ---
  # These plots show how the model output changes as each parameter is varied
  # one at a time, helping to identify which parameters have the strongest influence.
  sa_plots <- PEcAn.uncertainty::plot_sensitivities(
    sensitivity.results[[pft$name]]$sensitivity.output
  )
  print(do.call(gridExtra::grid.arrange, c(sa_plots, ncol = floor(sqrt(length(sa_plots))))))

  # --- 3. Generate Variance Decomposition Plots ---
  # These plots break down the total output variance into contributions from
  # each parameter, highlighting the most important sources of uncertainty.
  vd_plots <- PEcAn.uncertainty::plot_variance_decomposition(
    sensitivity.results[[pft$name]]$variance.decomposition.output
  )
  print(do.call(gridExtra::grid.arrange, c(vd_plots, ncol = 4)))
} else {
  PEcAn.logger::logger.warn("Could not find sensitivity results file:", sens_file)
}
```

# Customizing Ensemble Analysis Parameters (Optional)

## (Optional) Use this section only if you want to override the default ensemble analysis parameters. Skip if defaults are sufficient.
> **Important:** If you modify the ensemble analysis parameters in this section, re-run Section @sec-run-uncertainty and then Section @sec-visualize to regenerate outputs and plots.

```{r customize-ensemble}
# Set the number of ensemble members (model runs)
settings$ensemble$size <- 50

# Specify the variable(s) to be analyzed in the ensemble
# Single variable:
settings$ensemble$variable <- "NEE"
# Multiple variables (uncomment to use):
# settings$ensemble$variable <- c("NEE", "NPP", "LAI")

# Choose the method for sampling the parameter space (options: "uniform", "lhc", "halton", "sobol", "torus")
settings$ensemble$samplingspace$parameters$method <- "halton"
```


# Customizing Sensitivity Analysis Parameters (Optional)

## (Optional) Use this section only if you want to override the default sensitivity analysis parameters. Skip if defaults are sufficient.
> **Important:** If you modify the sensitivity analysis parameters in this section, re-run Section @sec-run-uncertainty and then Section @sec-visualize to regenerate outputs and plots.

```{r customize-sensitivity}
# Set the quantiles (in standard deviations) for parameter distribution in sensitivity analysis
settings$sensitivity.analysis$quantiles$sigma <- c(-2, -1, 1, 2)

# Specify the variable to be analyzed in sensitivity analysis
settings$sensitivity.analysis$variable <- "NEE"
```

# Extract Model Results and Prepare for Analysis

After the model simulation completes, we need to extract the results and prepare them for analysis. This involves:

1. Reading the run ID
2. Setting up output paths
3. Defining time period
4. Loading model output

```{r get-plot-vars}
runid <- as.character(read.table(paste(settings$outdir, "/run/", "runs.txt", sep = ""))[1, 1]) # Note: if you are using an xml from a run with multiple ensembles this line will provide only the first run id
# You can change [1,1] to [10,1], [5,1], etc. to select different run IDs from runs.txt
# For example: [10,1] selects the 10th run ID, [5,1] selects the 5th run ID
outdir <- paste(settings$outdir, "/out/", runid, sep = "")
start.year <- as.numeric(lubridate::year(settings$run$start.date))
end.year <- as.numeric(lubridate::year(settings$run$end.date))
model_output <- PEcAn.utils::read.output(
  runid,
  outdir,
  start.year,
  end.year,
  variables = NULL,
  dataframe = TRUE,
  verbose = FALSE
)
available_vars <- names(model_output)[!names(model_output) %in% c("posix", "time_bounds")]
```

# Display Available Model Variables

This section shows all the variables that are available in the model output. These variables represent different ecosystem processes and states that the model has simulated.

```{r show-variables}
vars_df <- PEcAn.utils::standard_vars |>
  dplyr::select(
    Variable = Variable.Name,
    Description = Long.name
  ) |>
  dplyr::filter(Variable %in% available_vars) |>
  # TODO: add year to PEcAn.utils::standard vars
  dplyr::bind_rows(
    dplyr::tibble(
      Variable = "year",
      Description = "Year"
    )
  )

vars_df$Description[is.na(vars_df$Description)] <- "(No description available)"
knitr::kable(vars_df, caption = "Model Output Variables and Descriptions")
```

# Visualize Model Results

This section provides examples of how to create time series plots of different model variables. The examples cover various ecosystem processes including carbon fluxes, carbon pools, water variables, and structural variables like Leaf Area Index (LAI).

## Plot Carbon Fluxes

```{r plot-carbon-fluxes}
# Plot Gross Primary Productivity (GPP) and Net Primary Productivity (NPP)
plot(model_output$posix, model_output$GPP,
  type = "l",
  col = "green",
  xlab = "Date",
  ylab = "Carbon Flux (kg C m-2 s-1)",
  main = paste("Carbon Fluxes Over Time - PEcAn", runid)
)
lines(model_output$posix, model_output$NPP, col = "blue")
legend("topright", legend = c("GPP", "NPP"), col = c("green", "blue"), lty = 1)
```

## Plot Carbon Pools

```{r plot-carbon-pools}
# Plot Total Live Biomass and Total Soil Carbon
plot(model_output$posix, model_output$TotLivBiom,
  type = "l",
  col = "darkgreen",
  xlab = "Date",
  ylab = "Carbon Pool (kg C m-2)",
  main = paste("Carbon Pools Over Time - PEcAn", runid)
)
lines(model_output$posix, model_output$TotSoilCarb, col = "brown")
legend("topright", legend = c("Total Live Biomass", "Total Soil Carbon"), col = c("darkgreen", "brown"), lty = 1)
```

## Plot Water Variables

```{r plot-water-variables}
# Plot Soil Moisture and Snow Water Equivalent
plot(model_output$posix, model_output$SoilMoist,
  type = "l",
  col = "blue",
  xlab = "Date",
  ylab = "Soil Moisture (kg m-2)",
  main = paste("Soil Moisture Over Time - PEcAn", runid)
)
lines(model_output$posix, model_output$SWE, col = "lightblue")
legend("topright", legend = c("Soil Moisture", "Snow Water Equivalent"), col = c("blue", "lightblue"), lty = 1)
```

## Plot LAI and Biomass

```{r plot-lai-biomass}
# Plot Leaf Area Index (LAI) and Above Ground Wood
plot(model_output$posix, model_output$LAI,
  type = "l",
  col = "darkgreen",
  xlab = "Date",
  ylab = "LAI (m2 m-2)",
  main = paste("Leaf Area Index Over Time - PEcAn", runid)
)
lines(model_output$posix, model_output$AbvGrndWood, col = "brown")
legend("topright", legend = c("LAI", "Above Ground Wood"), col = c("darkgreen", "brown"), lty = 1)
```

# Your turn

Try visualizing other variables in the model output! Use the `available_vars` object to see what variables are available and create your own time series plots. You can also customize the plots with different colors, labels, and titles.

Try editing the `pecan.xml` file. Give it a new name and update the **settings_path** variable at the beginning of this Demo to point to the new file. See how the changes affect the model output!

Inspect `sobol_indices` and identify which uncertainty class has the largest first-order contribution to variance, and which has the largest total-order contribution. Differences between first-order and total-order indices indicate the importance of interactions. <!-- TODO: add Sobol GSA visualization and interpretation example -->

# Conclusion

This notebook demonstrated how to set up, run, and analyze a PEcAn ecosystem model workflow programmatically. You can now modify parameters, try different models, or extend the analysis as needed.

# Further Exploration

The [next set of tutorials](#demo-table) will focus on the process of data assimilation and parameter estimation. The next two steps are in “.Rmd” files which can be viewed online. 

**Assimilation 'by hand'**

[Explore](https://github.com/PecanProject/pecan/blob/main/documentation/tutorials/sensitivity/PEcAn_sensitivity_tutorial_v1.0.Rmd) how model error changes as a function of parameter value (i.e. data assimilation ‘by hand’)


**MCMC Concepts**
[Explore](https://github.com/PecanProject/pecan/blob/main/documentation/tutorials/MCMC/MCMC_Concepts.Rmd) Bayesian MCMC concepts using the photosynthesis module

**More info about tools, analyses, and specific tasks…**

Additional information about specific tasks (adding sites, models, data; software updates; etc.) and analyses (e.g. data assimilation) can be found in the PEcAn [documentation](https://pecanproject.github.io/pecan-documentation/)

If you encounter a problem with PEcAn that’s not covered in the documentation, or if PEcAn is missing functionality you need, please search [known bugs and issues](https://github.com/PecanProject/pecan/issues?q=), submit a [bug report](https://github.com/PecanProject/pecan/issues/new/choose), or ask a question in our [chat room](https://join.slack.com/t/pecanproject/shared_invite/enQtMzkyODUyMjQyNTgzLWEzOTM1ZjhmYWUxNzYwYzkxMWVlODAyZWQwYjliYzA0MDA0MjE4YmMyOTFhMjYyMjYzN2FjODE4N2Y4YWFhZmQ).

# Clean Up Workflow Output (Optional)

If you want to remove all files and directories created by this workflow and start fresh, you can run the following code. This will delete the entire output directory specified in your settings. **Use with caution!**

```{r cleanup}
# WARNING: This will permanently delete all workflow output files!
# Uncomment the line below to enable cleanup.
# fs::dir_delete(settings$outdir)
```


# Session Information


### PEcAn package versions.

```{r version-info}
PEcAn.all::pecan_version()
```

### R session information:

```{r session-info}
sessionInfo()
```

# References

Dietze, M., 2017. Ecological Forecasting. Princeton University Press.

LeBauer, D. S., Wang, D., Richter, K. T., Davidson, C. C., and Dietze, M. C. 2013. Facilitating feedbacks between field measurements and ecosystem models. *Ecological Monographs* 83(2):133-154.

Puy, A., Lo Piano, S., Saltelli, A., and Levin, S. A. 2022. sensobol: An R Package to Compute Variance-Based Sensitivity Indices. *Journal of Statistical Software* 102(5):1-37. https://doi.org/10.18637/jss.v102.i05

Saltelli, A., Ratto, M., Andres, T., Campolongo, F., Cariboni, J., Gatelli, D. et al. 2008. *Global Sensitivity Analysis: The Primer*. New York: John Wiley & Sons.
